---
title: "Itasca_annual_summaries"
author: "Sara DeLaurentis"
date: "4/11/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

  
library(lubridate)
library(stringr)
library(tidyverse)
library(cowplot)
library(stringr)
library(dplyr)
library(glue) #from dplyr


```



##Create a dataframe to enable selection of .csv files.

```{r }
#create list of .csv files
m <- list.files("C:/Users/sbaue/coding/R_TEMPRY/Itasca_project_19-21/iButtons", pattern = "*.csv")
m

file_df <- data.frame(paths = m, name = str_replace(m, ".csv", ""), full_paths = list.files("C:/Users/sbaue/coding/R_TEMPRY/Itasca_project_19-21/iButtons", pattern = "*.csv", full.names = TRUE))

file_df

meta <- file_df$name %>%   #take the name column 
   str_split_fixed("_", n=5) %>% #split the strings, with _ as separator.
   as.data.frame() #COERCE
  colnames(meta) <- c("site", "rep", "position", "buttonID", "season") #
  meta$CLEAN = NULL 
  
  
data.list.df <- cbind(file_df, meta)
  #  temp_df <- temp_df[, c(6,1,2,3,4,5,7,8)] # column re-ordering if necessary
data.list.df <- data.list.df[order(data.list.df$site, data.list.df$rep, data.list.df$position, data.list.df$season), ] #for checking stuff

data.list.df <- data.list.df[order(data.list.df$name),]

view(data.list.df)

```


### THIS IS THE CODE YOU'VE BEEN LOOKING FOR !!! ###
### VERY COOOlll lists ###

#What this does:

#1. Takes the dataframe we created from the names of the selected files, turns each ROW into ONE list element with "site", "rep", etc. as named elements contained inside.

#2. Uses lapply to read in the .csv files and attaches metadata.


```{r }

files.meta.list <- list()                   # Create empty list

for(i in 1:nrow(data.list.df)) {   #for(each variable in the sequence "1 through the number of rows in C2A"):
                          
  files.meta.list[[i]] <- data.list.df[i , ]  #create an element in files.meta.list from each row found in C2A
}
#print(files.meta.list)

#Step 2: 

csv_LIST <- lapply(seq_along(files.meta.list), function(i){   
  df <- read.csv(files.meta.list[[i]]$full_paths, header=TRUE, sep = ",") #create "df" by reading the .csv given by full_paths.
  df$site <- files.meta.list[[i]]$site    #next few rows are the addition of metadata columns based on the named element in the list. 
  df$rep <- files.meta.list[[i]]$rep
  df$position <- files.meta.list[[i]]$position
  df$buttonID <- files.meta.list[[i]]$buttonID
  df$season <- files.meta.list[[i]]$season
  #Converting the date.time column. There may be multiple date formats that need to be dealt with. 
 #print(df[1, "date.time"])
  df$date.time <- mdy_hms(df$date.time, tz = "America/Chicago", truncated = 3) #Get R to recognize your current date format
  df$date.time <- as.POSIXct((round(df$date.time, "mins")), tz = "America/Chicago") #Round off the seconds (turns into POSIXlt), re-convert to POSIXct.
 # print(df[1, "date.time"]) # for checks
 # print(df[4, "date.time"]) #for checks
  
  df #VERY IMPORTANT. I don't know why, but without this call for the whole dataframe, you get left with just the last column you created.
  })

finalTable <- do.call(rbind, csv_LIST) ##JOIN TOGETHER

str(finalTable)
head(finalTable)
view(finalTable)


#output problem table if needed.
#problems <- finalTable %>% filter(is.na(date.time))
#view(problems)
#view(problems[problems$buttonID == "i45",])
#view(finalTable[finalTable$buttonID == "i101" & finalTable$season == "2021",])


#ggplot stuff:

#ggplot(finalTable, aes(x=as.POSIXct(date.time, format = "%m/%d/%Y %H:%M"), y=value, group = position, color = position)) + 
#  geom_line() + 
#  xlab("Date/Time") + 
#  ylab("Temperature in C")


```
## ANNOTATED COPY ##


### THIS IS THE CODE YOU'VE BEEN LOOKING FOR !!! ###
### VERY COOOlll lists ###

#What this does:

#1. Takes the dataframe we created from the names of the selected files, turns each ROW into ONE list element with "site", "rep", etc. as named elements contained inside.

#2. Uses lapply to read in the .csv files and attaches metadata.


```{r }

#Step 1: Turn each row of metadata into a list element with filenames, site, rep, etc. as named elements within.


files.meta.list <- list()                   # Create empty list


#Next, we can use a for-loop to store each row of our data frame as list element.

for(i in 1:nrow(data.list.df)) {   #for(each variable in the sequence "1 through the number of rows in C2A"):
                          
  files.meta.list[[i]] <- data.list.df[i , ]  #create an element in files.meta.list from each row found in C2A
}

#check it out: each element of the list (each [[x]]) contains the filenames and metadata, and they've maintained their column names as labels, so we can call out that named element when we want.

print(files.meta.list)


#Step 2: Use lapply to perform the following on each list element:
#Read .csv file given by the full_paths list element.
#Assign metadata columns based on the other named elements.
#Then, rbind all of those dataframes together for your finalTable

csv_LIST <- lapply(seq_along(files.meta.list), function(i){   #perform the given function along this list, each element in the list is the input.
  df <- read.csv(files.meta.list[[i]]$full_paths, header=TRUE, sep = ",") #create "df" by reading the .csv given by full_paths.
  df$site <- files.meta.list[[i]]$site    #next few rows are the addition of metadata columns based on the named element in the list. "Make a site column using the thing in the list at [[i]], that's labeled "site"
  df$rep <- files.meta.list[[i]]$rep
  df$position <- files.meta.list[[i]]$position
  df$buttonID <- files.meta.list[[i]]$buttonID
  df$season <- files.meta.list[[i]]$season
   #Converting the date.time column. There may be multiple date formats that need to be dealt with. 
 #print(df[1, "date.time"])
  df$date.time <- mdy_hms(df$date.time, tz = "America/Chicago", truncated = 3) #Get R to recognize your current date format
  df$date.time <- as.POSIXct((round(df$date.time, "mins")), tz = "America/Chicago") #Round off the seconds (turns into POSIXlt), re-convert to POSIXct.
 # print(df[1, "date.time"]) # for checks
 # print(df[4, "date.time"]) #for checks
  df #VERY IMPORTANT. I don't know why, but without this call for the whole dataframe, you get left with just the last column you created.
  
  })

#finalTable <- do.call(rbind, csv_LIST) ##JOIN TOGETHER
#view(finalTable)


#ggplot(finalTable, aes(x=as.POSIXct(date.time, format = "%m/%d/%Y %H:%M"), y=value, group = position, color = position)) + 
#  geom_line() + 
#  xlab("Date/Time") + 
#  ylab("Temperature in C")


```


## Mean Annual Temperature Summaries ##

```{r }
#data <- finalTable  #use this if you generated the finalTable using the code above.

data <- read_csv("ALL.csv")  #use this if you have a working table saved as ALL.csv.
head(data)
head(data$date.time)
data[30000,]
class(data[4,"date.time"])


#For the year 2020, Jan 1 - Dec 31.
START <- as.POSIXct("2020-01-01", tz = "")
BREAK <- as.POSIXct("2020-08-25", tz = "")
RESUME <- as.POSIXct("2020-09-30", tz = "")
END <- as.POSIXct("2020-12-31", tz = "")


data <- data %>% 
  subset.data.frame(date.time >= START & date.time <= END)  #subset the whole df, including the break interval
data <- data %>% 
  subset.data.frame(date.time <= BREAK | date.time >= RESUME) #then, cut out the break interval

#annual.sums.df <- data %>%                    
 #  subset.data.frame(date.time <= BREAK || date.time >= RESUME)
#filter(date.time >= START & date.time <= END) %>% 
 # group_by(site, rep, position, season) %>% 
 # summarise(meantemp = mean(value))

view(data)
str(data)
check <- data %>% 
  filter(date.time > as.POSIXct("2020-12-28", tz = ""))

view(check)

#annual.sums.df <- data %>%                    
 #  subset.data.frame(date.time <= BREAK || date.time >= RESUME)
#filter(date.time >= START & date.time <= END) %>% 
 # group_by(site, rep, position, season) %>% 
 # summarise(meantemp = mean(value))

```


##Annual summaries for September 2020 - September 2021

```{r}
#Annual summaries for October 2020 - September 2021: MEAN, MAX, MIN

data <- read_csv("ALL.csv")

START <- as.POSIXct("2020-10-01", tz = "")
#BREAK <- as.POSIXct("2020-08-25", tz = "")
#RESUME <- as.POSIXct("2020-09-30", tz = "")
END <- as.POSIXct("2021-09-30", tz = "")


data <- data %>% 
  subset.data.frame(date.time >= START & date.time <= END)  #subset the whole df, including the break interval

annual.sums.df.OCT <- data %>%
  group_by(site, rep, position) %>% 
  summarise(meantemp = mean(value), 
            min = min(value),
            min_temp_time = date(date.time[which.min(x = value)]),
            max = max(value),
            max_temp_time = date(date.time[which.max(x = value)]),
            length = length(date.time)) 

#how many values are in each group? Use this column to weed out sensors with only a few cells of data.

annual.sums.df.OCT <- annual.sums.df.OCT %>% 
  subset.data.frame(length > 500) #only select rows where the length column value is greater than 500.
annual.sums.df.OCT$length = NULL # delete length column       


view(annual.sums.df.OCT)
write_csv(annual.sums.df.OCT, "annual_sums_OCT_1.csv") 
```


```{r}
plot.data <- data %>% 
  group_by(site, position) 
plot.data <-plot.data %>% 
  subset.data.frame(length(date.time) >500) %>% 
  group_by(site, position)


str(plot.data)

means <- 

as.factor(plot.data$site)
#as.factor(plot.data$rep)
#as.factor(plot.data$position)
#as.factor(plot.data$buttonID)

myplot <- ggplot(data = plot.data, aes(site, mean(value)), fill = "blue") + 
  geom_col(position = "dodge")
myplot <- myplot + facet_grid(rows = vars(position))
myplot <- myplot + xlab("Site") + ylab("Temperature in C")
  
print(myplot)



## New graph trial

plot.data <- annual.sums.df.OCT
view(plot.data)
plot.data$site <- as.factor(plot.data$site)

plot.means <- annual.sums.df.OCT %>% 
  group_by(site, position) %>% 
  summarise(mean = mean(meantemp))
view(plot.means)

myplot2 <- ggplot(plot.data, aes(x = site, y = meantemp)) +
  geom_point(size = .6) + 
  geom_bar(data = plot.means, stat = "identity")
myplot2 <- myplot2 + facet_grid(rows = vars(position))

ggsave("plot_test.png", width = 5, height = 6) 
#print(myplot2)

```

```{r}
plot.data.z <- data %>% 
  group_by(site, rep, position) %>% 
  nest()
str(plot.data.z)



```

